---
title: "Moisture Analysis Hirsikorpi Field"
output:
  html_document:
  df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r config_knitr_options, echo=FALSE}
# Speeds up knitr by caching R output. 
# Clear Knitr Cache in the Knit menu empties the cache 
knitr::opts_chunk$set(cache = TRUE)
# echo=FALSE hides the code from Kint output
knitr::opts_chunk$set(echo = FALSE) 
# echo=FALSE hides the messages from Kint output
knitr::opts_chunk$set(message = FALSE)

knitr::opts_chunk$set(comment=NA)

```

```{r load_libraries}
library(psych)

# library
library(ggplot2)
library(tidyverse)
library(GGally)
library(lubridate)
```
# Overview of Data

## Weather Station Data
```{r load_Hirsikorpi_weather_data}
# The xls files are read and data is cleaned up in read_weather_station_Hirsikorpi.R file
# This file is located in the R subdirectory of the project folder
# If the data format of the data needs to be changed it is better to edit 
# read_weather_station_Hirsikorpi.R file and then rerun in with the following command

# source('./R/read_weather_station_Hirsikorpi.R')

# Load the Hirsikorpi weather station data from the R file located in Data/R/ folder
# loads Hirsikorpi_weather dataframe in the the global environment
load("../Data/R/Hirsikorpi_weather.R")

# Displays the first 6 rows of Hirsikorpi_weather
head(Hirsikorpi_weather)

data_Hirsikorpi_weather <- Hirsikorpi_weather
```
* Weather measurement on the following dates:
```{r}
unique(date(Hirsikorpi_weather$timestamp))
```


## Moisture Data

```{r load_Hirsikorpi_moisture_data}
# The cvs file is read and data is cleaned up in read_moisture_Hirsikorpi.R file
# This file is located in the R subdirectory of the project folder
# If the data format of the data needs to be changed it is better to edit 
# read_moisture_Hirsikorpi.R file and then rerun in with the following command

# source('./R/read_moisture_Hirsikorpi.R')


# Load the Hirsikorpi moisture data from the R file located in Data/R/ folder
# loads Hirsikorpi_moisture dataframe in the the global environment
load("../Data/R/Hirsikorpi_moisture.R")

# Displays the first 6 rows of Hirsikorpi_moisture
head(Hirsikorpi_moisture)
```
* Moisture measurement on the following dates:
```{r}
unique(date(Hirsikorpi_moisture$timestamp))
```

## Ground Water Data
```{r load_ground_water_data}
# The cvs file is read and data is cleaned up in read_ground_water_Hirsikorpi.R file
# This file is located in the R subdirectory of the project folder
# If the data format of the data needs to be changed it is better to edit 
# read_ground_water_Hirsikorpi.R file and then rerun in with the following command

# source('./R/read_ground_water_Hirsikorpi.R')


# Load the Hirsikorpi ground water data from the R file located in Data/R/ folder
# loads Hirsikorpi_ground_water dataframe in the the global environment
load("../Data/R/Hirsikorpi_ground_water.R")

head(Hirsikorpi_ground_water)
```
* Ground water measurement on the following dates:
```{r}
unique(date(Hirsikorpi_ground_water$timestamp))
```


```{r load_Hirsikorpi_combined}

# Hirsikorpi_moisture, Hirsikorpi_ground_water and Hirsikorpi_weather are joined in join_data_Hirsikorpi.R file
# This file is located in the R subdirectory of the project folder
# If the joining needs t be changed it is better to edut 
# join_data_Hirsikorpi.R file and then rerun in with the following command

# source('./R/join_data_Hirsikorpi.R')


# Load the Hirsikorpi_r data from the R file located in Data/R/ folder
# loads Hirsikorpi_combined dataframe in the the global environment
load("../Data/R/Hirsikorpi_combined.R")

```


## Combined Data

### Descriptive Statistics 

```{r, exploreData}
head(Hirsikorpi_combined)

# Calculate the descriptive statistics for the columns of Hirsikorpi_combined
# timestamp and time_diff columns are excluded.  
Hirsikorpi_combined %>% 
  select(-timestamp, -timestamp_weather, -timestamp_ground_water, -time_diff_weather, -time_diff_ground_water) %>% 
  describe()

```

### Missing Data

Identification of moisture measurements for which weather station and ground water data are not available

#### Missing Weather Station Data

```{r}
ggplot(Hirsikorpi_combined, aes(x=timestamp, y=as.numeric(time_diff_weather))) + 
  theme_minimal() + 
  geom_point() +
  ylab("Time difference in minitues")
```


Data points for which weather data measurements are not within a certain time frame are removed from the dataset. 

* Initial number of data points: `r nrow(Hirsikorpi_combined)`

```{r}
# The cut off value is set to 5 mins.
# This value (5) is chosen arbitrarily. It can be changed to a more meaningfull value
time_frame_for_weather_data <- 5  

number_of_rows <- nrow(Hirsikorpi_combined)
# Get total number of rows in the Hirsikorpi_combined data
column_names <- names(Hirsikorpi_combined) 
Hirsikorpi_combined_filtered <- Hirsikorpi_combined %>% 
  filter(abs(time_diff_weather) > 5)
```

* Number of data points removed when threshold for time frame is set to `r time_frame_for_weather_data`: `r number_of_rows - nrow(Hirsikorpi_combined)`
* Remanining number of data points: `r nrow(Hirsikorpi_combined)`


#### Missing Ground Water Data

```{r}
ggplot(Hirsikorpi_combined, aes(x=timestamp, y=as.numeric(time_diff_ground_water))) + 
  theme_minimal() + 
  geom_point() +
  ylab("Time difference in minitues")
```

For all moisture measurements the time between ground water measurement is less than 2 minutes. None of the data is removed.

### Descriptive Statistics for Filtered Data

```{r}
# Calculate the descriptive statistics for the columns of Hirsikorpi_combined
# timestamp and time_diff columns are excluded.  
Hirsikorpi_combined_filtered %>% 
  select(-timestamp, -timestamp_weather, -timestamp_ground_water, -time_diff_weather, -time_diff_ground_water) %>% 
  describe()
```

# Plots
## Histogram
```{r, histogram}
# Basic histogram
ggplot(Hirsikorpi_combined_filtered, aes(x=moisture1)) + geom_histogram(binwidth = 0.005) + theme_minimal()
ggplot(Hirsikorpi_combined_filtered, aes(x=moisture2)) + geom_histogram(binwidth = 0.005) + theme_minimal()


# Custom Binning. I can just give the size of the bin
ggplot(Hirsikorpi_combined_filtered, aes(x=moisture3)) + geom_histogram(binwidth = 0.005)
```
## Scatterplots
```{r, scatterplots}
# basic scatterplot
ggplot(Hirsikorpi_combined_filtered, aes(x=timestamp, y=moisture3)) + 
  geom_point()
```

```{r, plotXls_Hirsikorpi}
# Custom Binning. I can just give the size of the bin
ggplot(Hirsikorpi_combined_filtered, aes(x=temperature)) + geom_histogram()


```

```{r plotPairs}

ggpairs(Hirsikorpi_combined_filtered, columns=c("temperature", "wind_gust", "humidity", "moisture1", "moisture2", "moisture3"))

```


```{r}
ggpairs(Hirsikorpi_combined_filtered, columns=c("MA_2_1", "MA_2_2", "MA_2_3", "MA_2_4", "humidity", "moisture1", "moisture2", "moisture3"))
```

```{r -scatterpot ind variables timestamp/temperature}
ggplot(data = Hirsikorpi_weather) + 
  geom_point(mapping = aes(x = timestamp, y = temperature, color = humidity))
```

```{r -geom_smooth temperature/timestamp}

ggplot(data = Hirsikorpi_weather) + 
  geom_smooth(mapping = aes(x = timestamp, y = temperature))
```


```{r -scatterpot ind variables temperature/humidity}
ggplot(data = Hirsikorpi_weather) + 
  geom_point(mapping = aes(x = temperature, y = humidity, color = timestamp))
```

```{r -geom_smooth humidity/temperature}
ggplot(data = Hirsikorpi_weather) + 
  geom_smooth(mapping = aes(x = temperature, y = humidity))
```

```{r -scatterpot ind variables timestamp/humidity}
ggplot(data = Hirsikorpi_weather) + 
  geom_point(mapping = aes(x = timestamp, y = humidity, color= temperature))
```

```{r -geom_smooth humidity/timestamp}
ggplot(data = Hirsikorpi_weather) + 
  geom_smooth(mapping = aes(x = timestamp, y = humidity))
```


```{r -scatterpot ind variables waterheight1/timestamp}
ggplot(data = Hirsikorpi_ground_water) + 
  geom_point(mapping = aes(x = timestamp, y = MA_2_1), position = "jitter" )
```

```{r -geom_smooth ground_water_1/timestamp}
ggplot(data = Hirsikorpi_ground_water) + 
  geom_smooth(mapping = aes(x = timestamp, y = MA_2_1))
```


```{r -scatterpot ind variables waterheight2/timestamp}
ggplot(data = Hirsikorpi_ground_water) + 
  geom_point(mapping = aes(x = timestamp, y = MA_2_2), position = "jitter" )
```

```{r -geom_smooth ground_water_2/timestamp}
ggplot(data = Hirsikorpi_ground_water) + 
  geom_smooth(mapping = aes(x =timestamp , y = MA_2_2))
```

```{r -scatterpot ind variables timestamp/waterheight3}
ggplot(data = Hirsikorpi_ground_water) + 
  geom_point(mapping = aes(x = timestamp, y = MA_2_3), position = "jitter" )
```

```{r -geom_smooth ground_water_3/timestamp}
ggplot(data = Hirsikorpi_ground_water) + 
  geom_smooth(mapping = aes(x = timestamp, y = MA_2_3))
```

```{r -scatterpot ind variables waterheight1/waterheight2}
ggplot(data = Hirsikorpi_ground_water) + 
  geom_point(mapping = aes(x = MA_2_1, y = MA_2_2, color = timestamp), position = "jitter" )
```

```{r -scatterpot ind variables waterheight1/waterheight3}
ggplot(data = Hirsikorpi_ground_water) + 
  geom_point(mapping = aes(x = MA_2_1, y = MA_2_3, color = timestamp), position = "jitter" )
```
```{r -scatterpot ind variables waterheight2/waterheight3}
ggplot(data = Hirsikorpi_ground_water) + 
  geom_point(mapping = aes(x = MA_2_2, y = MA_2_3, color = timestamp), position = "jitter" )
```

```{r -scatterpot ind variables temperature/waterheight1}
ggplot(data = Hirsikorpi_ground_water) + 
  geom_point(mapping = aes(x = MA_2_2, y = MA_2_3, color = timestamp), position = "jitter" )
```
```{r -scatterpot ind variables temperature/waterheight2}
ggplot(data = Hirsikorpi_combined_filtered) + 
  geom_point(mapping = aes(x = temperature, y = MA_2_2, color = timestamp), position = "jitter" )
```

```{r -scatterpot ind variables temperature/waterheight3}
ggplot(data = Hirsikorpi_combined_filtered) + 
  geom_point(mapping = aes(x = temperature, y = MA_2_3, color = timestamp), position = "jitter" )
```


```{r -temperature/humidity}
ggplot(data = Hirsikorpi_weather) + 
  geom_smooth(mapping = aes(x = temperature, y = humidity))
```



